---
output:
  pdf_document: default
  html_document: default
---
# Juno-typing {#juno-typing}

<style>
body {
text-align: justify}
</style>

```{r setup_juno, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = FALSE)
```

```{r echo = FALSE, eval = TRUE}

# Parameters
parameters <- list("pipeline_name" = "Juno-typing")

```

The goal of this pipeline is to perform bacterial typing (7-locus MLST and serotyping). It takes 2 types of files per sample as input:  

1. Two '.fastq' files (paired-end sequencing) derived from short-read sequencing. They should be already filtered and trimmed (for instance, with the Juno-pipeline).  

2. An assembly from the same sample in the form of a single '.fasta' file.  

Importantly, the Juno-typing pipeline works directly on output generated from the Juno-assembly pipeline. 

The Juno-typing pipeline will then make the following analyses:  

- 7-locus MLST (using the [MLST tool](https://bitbucket.org/genomicepidemiology/mlst/src/master/) from the CGE group).  

- Serotyping Salmonella samples (using the [SeqSero2](https://aem.asm.org/content/85/23/e01746-19) tool).  

- Serotyping E. coli samples (using the [SerotypeFinder](https://bitbucket.org/genomicepidemiology/serotypefinder/src/master/) tool).  

<span style="color:red;">Disclaimer!!</span> Importantly, the genus and species are automatically detected from the processed reads (using `kmerFinder`) and the schema to use for the 7-locus MLST is chosen accordingly. The results of `kmerfinder` are also included in the output. However, <span style="font-weight:bold;">this pipeline is NOT meant for bacteria identification</span> and it has not been tested for it. Even if similar species would be misidentified, they would probably share the same MLST7 scheme. Following that thought, the correct choice of scheme has been tested, but not the correct species identification.  

## Handbook

### Requirements and preparation

See the [General Instructions for all pipelines](#general-instructions) first.  

- This pipeline needs two fastq file (R1 and R2) and an assembly (.fasta) files per sample. The fastq files should have been trimmed and filtered to remove low quality reads/bases. You could use the [Juno-assembly pipeline](#juno-assembly) for that. Moreover, that pipeline also provides the _de novo assembly_ for your samples and the output folder of the `Juno-assembly` pipeline can be used directly into the `Juno-typing` pipeline. If you, however, prefer to use any other tool for doing your assembly and trimming/filtering, make sure that the fastq files and fasta files have the same name (for instance, sample1_R1_001.fastq.gz, sample1_R1_001.fastq.gz and sample1.fasta). If that is not the case, the files may not be recoginzed as belonging to the same sample. Also, ALL THREE FILES SHOULD BE IN THE SAME FOLDER! If you have multiple samples, they should all be in the same input folder, NOT IN SUBFOLDERS. The only exception is if you use the `Juno-assembly` pipeline to pre-process your data. In that case, the pipeline will recognize the subfolders where the fastq files and the fasta files should be.


### Download the pipeline  

**YOU NEED TO DOWNLOAD THE PIPELINE ONCE OR EVERY TIME YOU WANT TO UPDATE IT**

Please follow the [instructions to download pipelines](#downloading-instructions) from the Juno team of the IDS-bioinformatics group. The `r parameters$pipeline_name` pipeline can be found in [this link](https://github.com/RIVM-bioinformatics/Juno-typing).  


### Start the analysis. Basics

1. Open a terminal. (Applications>terminal).  

2. Enter the folder of the pipeline using:  

```{bash}

cd /mnt/scratch_dir/<my_folder>/Juno-typing

```

3. Run the pipeline  

```{bash}
bash juno-typing -i /mnt/scratch_dir/<my_folder>/<my_data>/
```

Please read the section [What to expect while running a Juno pipeline](#what-to-expect)

See the section [General Troubleshooting](#general-troubleshooting) for any problems you may encounter. 

**Note:** Do not keep all your data (including results) on the scratch_dir partition. You are allowed to keep 400GB max and with sequencing data, this can get full quite fast.

### Output 

A folder called `out/`, inside the folder of the pipeline, will be created. This folder will contain all the results and logging files of your analysis. There will be one folder per tool (`kmerfinder`, `mlst7` and `serotype`). Please refer to the manuals of every tool to interpret the results. Each one of these folders, there should be a sub-folder per sample and, for the case of mlst7 and serotype, also a csv file collecting the results of all the samples together: a `serotype/salmonella_serotype_multireport.csv`, `serotype/ecoli_serotype_multireport.csv` and `mlst7/mlst7_multireport.csv`. Even if your samples are not _Salmonella_ or _E. coli_ you will get the multireport file, althought it will be empty. Although the results of `kmerfinder` are provided, these have not been validated. They are used only to choose the right scheme for the MLST and the right serotyper. If you would use the results of `kmerfinder` as a species identification tool, you do it under your own risk and you should be able to interpret correctly the results.  

*Note:* If you want your output to be stored in a folder with a different name or location, you can use the option `-o` ('o' from output) 

```{bash}
bash juno-typing -i /mnt/scratch_dir/<my_folder>/<my_data>/ -o /mnt/scratch_dir/<my_folder>/<my_results>/
```

Another very important output from the pipeline are the logging files and audit trail that contain information of the software versions used, the parameters used, the error messages, etc. They could be important for you if you want to publish or reproduce the analysis at a later timepoint and also to get help from the bioinformatics team if you were to run into trouble with the pipeline. Please read about these files [here](#general-output). 

### Troubleshooting for this pipeline

**Please read first the [General Troubleshooting](#general-troubleshooting) section!**

#### Other problems or failing rules  

The `r parameters$pipeline_name` pipeline is still in development which means that sometimes the process can fail.

Before contacting for help, try these two steps:

1. Re-run the pipeline again and see if the process continues. If it does, please keep re-running the pipeline until your analysis is finished or there is no longer progress. In this case, send an email after the pipeline is finished so I can troubleshoot the problem.

2. [Download](https://github.com/RIVM-bioinformatics/Juno-typing) the pipeline again and start from the beginning of this [handbook](#downloading-instructions). Sometimes there is an issue that has been resolved in newer versions of the pipeline.

If the pipeline still fails after these two steps, please inform me about the problem. Send an [e-mail](mailto:alejandra.hernandez.segura@rivm.nl) with the following content:

* The log and error files that can be found in the output folder
* The path to your input directory
* The path to where the pipeline is installed

**Note** I cannot help you without this information, if information is missing there will be an delay in troubleshooting the problem.