---
output:
  pdf_document: default
  html_document: default
---
# Triumph: Project report {#triumph_project}

<style>
body {
text-align: justify}
</style>

```{r setup_triumph_project_report, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = FALSE)
```

```{r echo = FALSE, eval = TRUE}

# Parameters
parameters <- list("pipeline_name" = "triumph_microbiome_project")

```

The code here generates a report that collects the results of multiple runs of the Triumph report. Note that in order to use it, all the individual runs should have been analyzed with the ["Triumph: Run pipeline"](#triumph_run)

## Handbook

### Requirements and preparation

See the [General Instructions for all pipelines](#general-instructions) first.  

- Make sure that your folder has the right name and structure. Your input folder should have one subfolder per different run. Each of this subfolders contains the results of the ["Triumph: Run pipeline"](#triumph_run) for that specific run. Please, do not rename any of the files produced by the ["Triumph: Run pipeline"](#triumph_run), otherwise they will not be found and cannot be included in the project report. Neither your folder nor your subfolders can have "rare" characters (only letters, numbers, dashes and underscores accepted). If your folder name contains different characters, it may not be recognized by the pipeline. IMPORTANT: Your folder name MUST NOT CONTAIN SPACES!!!!  

### Download the pipeline   

**YOU NEED TO DOWNLOAD THE PIPELINE ONCE OR EVERY TIME YOU WANT TO UPDATE IT**

Please follow the [instructions to download pipelines](#downloading-instructions) from the Juno team of the IDS-bioinformatics group. The `r parameters$pipeline_name` pipeline can be found in [this link](https://gitl01-int-p.rivm.nl/hernanda/triumph_project_report).  

### Start the analysis. Basics

1. Open the terminal.  

2. Enter the folder of the pipeline:  

```{bash}
cd /mnt/scratch_dir/<my_folder>/triumph_project_report-master/
```
* Exact name of downloaded folder might be slightly different depending on the version that was downloaded.

3. Run the pipeline  

```{bash}
bash create_triumph_report.sh -i /mnt/scratch_dir/<my_folder>/
```
* The input directory (in this case /mnt/scratch_dir/my_folder) should be a folder containing one subfolder per run you want to include in the project report. These subfolders should contain the results of running the [Triumph run pipeline](#triumph_run) using the `--metadata` option. Refer to the manual of that pipeline to see what that argument means.  

**Note 1:** If you are updating the project report because you added the analysis of an old run, you should first delete or rename the "old" report. This script might fail otherwise.

**Note 2:** THE PROJECT REPORT PIPELINE ONLY WORKS WHEN ALL RUNS COLLECTED WERE EVALUATED WITH THE ["Triumph: Run pipeline"](#triumph_run) USING A METADATA FILE!!! 

### Output 

A folder called `out/`, inside the folder of the pipeline, will be created. If you prefer to store your output in a different folder, you can use the `-o` argument to give another path.  

```{bash}
bash triumph_run_pipeline.sh -i /mnt/scratch_dir/<my_folder>/<my_data>/ -o /mnt/scratch_dir/<my_folder>/<my_results>/

```

If you chose another site for the output, then you will find the same results there. 

In both cases, the output folder will contain the Triumph_Project_report.html file that is the actual report and a Triumph_Project_Report.log that contains messages, warnings and errors from R, where you can find more info if something goes wrong and the report could not be produced. 


### Troubleshooting 

**Please read first the [General Troubleshooting](#general-troubleshooting) section!**


#### Other problems or failing rules  

The `r parameters$pipeline_name` pipeline is still in development which means that sometimes the process can fail. Make sure you ran the pipeline with the Triumph_dada2_run_pipeline and that you did not change the name of any file. Try to create the report again. If you are sure all the input files are correct and the report does not run try the two steps below before contacting for help:

1. Re-run the pipeline again and see if the process continues. If it does, please keep re-running the pipeline until your analysis is finished or there is no longer progress. In this case, send an email after the pipeline is finished so I can troubleshoot the problem.

2. [Download](https://gitl01-int-p.rivm.nl/hernanda/triumph_project_report) the pipeline again and start from the beginning of this [handbook](#triumph_project). Sometimes there is an issue that has been resolved in newer versions of the pipeline.

If the pipeline still fails after these two steps, please inform me about the problem. Send an [e-mail](mailto:alejandra.hernandez.segura@rivm.nl) with the following content:

* The log and error files that can be found in the output folder
* The path to your input directory
* The path to where the pipeline is installed

**Note** I cannot help you without this information, if information is missing there will be an delay in troubleshooting the problem.